\documentclass[12pt]{article}
\usepackage{lingmacros}
\usepackage{tree-dvips}
\usepackage{amsmath}
\usepackage{accents}
\newcommand{\ubar}[1]{\underaccent{\bar}{#1}}
\usepackage{hyperref}
\hypersetup{
    colorlinks=true,
    linkcolor=blue,
    filecolor=magenta,      
    urlcolor=cyan,
}
\usepackage{pgfplots}

\begin{document}

\title{SIE 550 HW 3 Meeting Notes}
\date{3/3/17}

\maketitle

Fork and contribute to these notes on: \url{https://github.com/eggsactly/SIE550}

\section*{Problem 1} 
We feel comfortable with problem 1, problem 3 looks like it will be hard because it is time variant. 

\section*{Problem 2} 
\subsection*{2.1}
Finding the eigenvalues, solve this equation:
$$det(A-I\lambda)=0$$
Look at the real part of all the eigenvalues, if all the values are less than zero the system is asymptotically stable.\\
If any values are greater than zero then the system is unstable. 

\subsection*{2.2}
What is it asking us to do?

\subsubsection*{2.2a}
Plug in 
$$P=\begin{bmatrix} p_{11} & p_{12} \\
p_{12} & p_{22}
\end{bmatrix}$$ 

and 

$$\ubar{x}=\begin{bmatrix} x_{1} \\
x_{2}
\end{bmatrix}$$ 

to

$$V(\ubar{x})=\ubar{x}^TP\ubar{x}$$

You will get a scalar out of that equation if you do the matrix multiplication. Then look at the two equations. 
$p_{11}>0$ 
and
$p_{11}p_{22}-p_{12}^2>0$ \\

For the second equation to be true, $p_{22}>0$ 

Then you would need to explain that $p_{11}x_{1}^2$ and $p_{22}x_{2}^2$ dominate $p_{12}x_1x_2$ in all values for $\ubar{x}$. 

\subsubsection*{2.2b}
Find P in $A^TP+PA=-Q$ where 
$A=\begin{bmatrix} 0 & -1 \\
1 & -1
\end{bmatrix}$
$Q=\begin{bmatrix} 1 & 0 \\
0 & 1
\end{bmatrix}$

You can still assume that $P=\begin{bmatrix} p_{11} & p_{12} \\
p_{12} & p_{22}
\end{bmatrix}$

At the end, you will have solved for $p_{11}$, $p_{12}$, $p_{22}$. Once you have solved for P, find the eigenvalues for P. If all the eigenvalues for P are positive then the matrix is positive definite. 

\section*{Problem 3}
\subsection*{3.1}
Look at page 20 on the week 3 lecture notes. 

We know that
$$A(t)=\begin{bmatrix} -1 & e^{2t} \\
0 & -1
\end{bmatrix}=\begin{cases}\dot{x_1}=-x_1+e^{2t}x_2\\ 
\dot{x_2}=-x_2\\ 
x(t_0)=[x_1(0), x_2(0)]^T\\
\end{cases}$$ 

Solve for each line:
$$\dot{x_2}=-x_2\implies x_2=e^{-t}x_2(0)$$
$$\dot{x_1}=-x_1+e^{2t}e_2=-x_1+e^{2t}e^{-t}x_2(0)$$
$$\dot{x_1}+x_1=e^tx_2(0)$$
$$\dot{x_1}=x_{1_H}+x_{1_P}$$
$$\dot{x_1}+x_1=0\implies x_{1_H}=-e^tx(0)$$ (Separation of variables)
$$x_{1_P}=Ae^t\implies Ae^t+Ae^t=-e^tx_2(0)\implies A=\frac{x_2(0)}{2}$$
$$x_1=-e^tx_1(0)+\frac{x_2(0)}{2}e^t$$
$$\bar{x}=\begin{bmatrix} -e^t & \frac{e^t}{2} \\
0 & e^{-t}
\end{bmatrix}
\begin{bmatrix}x_1(0)\\ 
x_2(0)\\ 
\end{bmatrix}$$ 

\subsection*{3.2}
We do not know what this question is asking as of 3/2/17.

\section*{Problem 4}
\subsection*{4.1}
To find the equilibrium points set $\dot{x_1}$ and $\dot{x_2}$ equal to zero. For each equation use the hint and solve for $x_1$ when $x_2=0$ and $x_2$ when $x_1=0$. Where both equations for $\dot{x_1}$ and $\dot{x_2}$ are zero is where your equilibrium points are. Don't worry about $f_1(x)$ and $f_2(x)$, because of the rules in the problem description these functions are zero when $x=0$.

\subsection*{4.2}
Remember, an invariant set is a set where a point that starts on the set (initial condition) stays in the set.
To show that M is an invariant set, notice that the set M is an ellipse. 

\begin{tikzpicture}
\begin{axis}[
    axis lines = left,
    xlabel = $x_1$,
    ylabel = {$x_2$},
]
%Below the red parabola is defined
\addplot [
    domain=-2.5:2.5, 
    samples=200, 
    color=red,
]
{sqrt(4-2*x^2)};
\addlegendentry{$x_1^2+2x_2^2=4$}
%Here the blue parabloa is defined
\addplot [
    domain=-2.5:2.5, 
    samples=200, 
    color=red,
]
{-sqrt(4-2*x^2)};

\end{axis}
\end{tikzpicture}

If you were to solve for one side, $x_1$ for example, you would get
$$x_1=\pm\sqrt{4-2x_2^2}$$

Plugging that into your equation for $\dot{x_1}$ and $\dot{x_2}$ you will find the entire right hand side of each equation will fall off. You will end up with:
$$\begin{bmatrix}\dot{x_1} \\ 
\dot{x_2}\end{bmatrix}=
\begin{cases}-4x_1^2x_2\\ 
2x_1^3
\end{cases}$$ 

Those two equations assume that you are plugging in a point on the ellipse. If you can find a Lyapunov function for that system, it will show that the system is stable, which means that it will go to the equilibrium points, which are on the set M. This means that when you start on the set M, you will stay in the set M, meaning M is invariant.  

\subsection*{4.3}
In this problem apply the Lyapunov function provided to the system. You'll find that the left hand side get's canceled out. The right hand side of each equation will be left, but it will always be less than or equal to zero if you apply the the rule $x_if(x_i)>0$.

\section*{Problem 7}
In lecture on 2/21/17 Furfaro said that the following problem shows up on the homework:
$$\begin{cases}\dot{x_1}=-x_1-g(t)x_2\\ 
\dot{x_2}=x_1-x_2\\
\end{cases}$$ 

$$0\leq g(t)\leq k$$
$$\dot{g(t)} \leq g(t), \forall t \leq 0$$

Choose a good Lyapunov function. 
$$v(\bar{x},t)=x_1^2+(1+g(t))x_2^2$$

1. is a good candidate\\
$$\alpha(||\bar{x}||)\leq v(\bar{x},t)\leq \beta(||\bar{x}||)$$

$$(x_1^2+x_2^2)\leq x_1^2+(1+g(t))x_2^2\leq x_1^2+(1+k)x_2^2$$

$$\leq(1+k)(x_1^2+x_2^2)$$

$$(x_1^2+x_2^2)\leq v(\bar{x},t)\leq (1+k)(x_1^2+x_2^2)$$
$$(x_1^2+x_2^2)\leq v(\bar{x},t)\leq (1+k)||\bar{x}||^2$$
$$\alpha(||\bar{x}||)=(x_1^2+x_2^2)$$
$$\beta=(||\bar{x}||)$$
$$c_1||\bar{x}||^2\leq v(\bar{x},t)\leq c_2||\bar{x}||^2$$
$c_1=1$, $c_2=1+k$
$$\dot{v}=2x_1\dot{x_1}+\dot{g(t)}x_2^2+(1+g(t))x_2\dot{x_2}$$
$$\dot{v}(\bar{x},t)=-2x_1^2+2x_1x_2-[2+2g(t)-\dot{g}(t)]x_2^2$$
$$\dot{v}=-[2x_1^2-2x_1x_2+[2+2g(t)-\dot{g}(t)]x_2^2]$$
$$\dot{v}=\bar{x}^TP\bar{x}=\begin{bmatrix} x_1 & x_2 \end{bmatrix}
\begin{bmatrix} p_{11} & p_{12} \\
p_{21} & p_{22} \end{bmatrix}\begin{bmatrix} x_1 \\ x_2 \end{bmatrix} = (p_{11}x_1+p_{12}x_2)x_1+(p_{12}x_1+p_{22}x_2)x_2$$
Because $\dot{g}\leq g$
$$2+2g(t)-\dot{g}(t)\geq 2+2g(t)=g(t)$$
$$2+2g(t)-\dot{g}(t)\geq 2+g(t)$$
$$0\leq g(t) \leq k$$
Conclusion:
$$\dot{v}\leq -2x_1^2+2x_2x_2+[\alpha+2g(t)-\dot{g}(t)]$$
$$\dot{v}\leq -2x_1^2+2x_1x_2-2x_2^2$$
$$\dot{v}\leq -2||\bar{x}||^2$$

\end{document}


